---
title: "basic regression"
author: "Jin Chen"
date: "9/4/2020"
output: html_document
---

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)

```
Three common steps in EDA
    1. Most crucially, looking at the raw data values.
    2. Computing summary statistics, such as means, medians, and interquartile ranges.
    3. Creating data visualizations.


```{r Numerical Outcome, Numerical Explanatory}
evals_ch5 <- evals %>%
  select(ID, score, bty_avg, age)

glimpse(evals_ch5)

evals_ch5 %>%
  sample_n(size = 5)

evals_ch5 %>%
  summarize(mean_bty_avg = mean(bty_avg), mean_score = mean(score),
            median_bty_avg = median(bty_avg), median_score = median(score))

evals_ch5 %>% 
  select(score, bty_avg) %>% 
  skim()

#getting correlation coefficient
evals_ch5 %>% 
  get_correlation(formula = score ~ bty_avg)

evals_ch5 %>% 
  summarize(correlation = cor(score, bty_avg))

#plot it
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", 
       y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")

#add a jitter
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")

#add a line of best fit
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```


(LC5.1) 
Conduct a new exploratory data analysis with the same outcome variable y being score but with age as the new explanatory variable x.

Remember, this involves three things:

Looking at the raw data values.
Computing summary statistics.
Creating data visualizations.

What can you say about the relationship between age and teaching scores based on this exploration?

```{r Learning Check}
evals_ch5 %>% 
  select(score, age)

evals_ch5 %>% 
  select(score, age) %>% 
  skim()

evals_ch5 %>% 
  summarize(correlation = cor(score, age))

evals_ch5 %>% 
  ggplot(aes(x = age, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Age",
       y = "Score",
       title = "Age and Score have a weak negative relationship")


```

```{r Simple Linear Regression}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch5)
# Get regression table:
get_regression_table(score_model)
```

```{r LC5.2}
score_model2 <- lm(score ~ age, data = evals_ch5)
get_regression_table(score_model2)

```

```{r Residuals}
regression_points <- get_regression_points(score_model)
regression_points
```

```{r LC5.3}
regression_points2 <- get_regression_points(score_model2)
regression_points2

```
Questions
Differences between continents: Are there significant differences in average life expectancy between the five populated continents of the world: Africa, the Americas, Asia, Europe, and Oceania?
Differences within continents: How does life expectancy vary within the world’s five continents? For example, is the spread of life expectancy among the countries of Africa larger than the spread of life expectancy among the countries of Asia?


```{r One categorical explanatory variable}

#we do an EDA first in 3 steps remember. 1. Look at the raw data 2. Look at summary stats and 3. Create a data viz
library(gapminder)
gapminder2007 <- gapminder %>%
  filter(year == 2007) %>%
  select(country, lifeExp, continent, gdpPercap)

#1
glimpse(gapminder2007)
gapminder2007 %>% sample_n(size = 5)

#2
gapminder2007 %>%
  select(lifeExp, continent) %>%
  skim()

#why is the mean less than the median?

#3 
ggplot(gapminder2007, aes(x = lifeExp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Life expectancy", y = "Number of countries",
       title = "Histogram of distribution of worldwide life expectancies")

#we see that there are a couple of countries with very low life exp. This brings down the mean, but the median is less sensitive to it.


ggplot(gapminder2007, aes(x = lifeExp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Life expectancy", 
       y = "Number of countries",
       title = "Histogram of distribution of worldwide life expectancies") +
  facet_wrap(~ continent, nrow = 2)

ggplot(gapminder2007, aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  labs(x = "Continent", y = "Life expectancy",
       title = "Life expectancy by continent")

#median and mean for each continent
lifeExp_by_continent <- gapminder2007 %>%
  group_by(continent) %>%
  summarize(median = median(lifeExp), 
            mean = mean(lifeExp))
lifeExp_by_continent

```


Conduct a new exploratory data analysis with the same explanatory variable x being continent but with gdpPercap as the new outcome variable y.

What can you say about the differences in GDP per capita between continents based on this exploration?
```{r LC5.4}
#1
glimpse(gapminder2007)
gapminder2007 %>% sample_n(size = 5)

#2
gapminder2007 %>%
  select(gdpPercap, continent) %>%
  skim()

#3
ggplot(gapminder2007, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  labs(x = "Continent", y = "Life expectancy",
       title = "Life expectancy by continent")

#median and mean for each continent
gdpPercap_by_continent <- gapminder2007 %>%
  group_by(continent) %>%
  summarize(median = median(gdpPercap), 
            mean = mean(gdpPercap))
gdpPercap_by_continent


```

```{r Linear Regression}
lifeExp_model <- lm(lifeExp ~ continent, data = gapminder2007)
get_regression_table(lifeExp_model)


```
With a categorical explanatory variable, the estimates are relative to a "baseline". The intercept here of 54.8 is Africa's mean lifeExp. The default is alphabetical, that's why we're using Africa. The other numbers correspond to the mean lifeExp of the other continents. For example, continentAmericas is the mean difference between Africa and Americas. 
Thus for every k number of categories, we get k minus 1 offsets. In this example we had 5 continents, returning 4 offsets. 


```{r LC5.5}
gdpPercap_model <- lm(gdpPercap ~ continent, data = gapminder2007)
get_regression_table(gdpPercap_model)


```

```{r Observed/fitted values and residuals}
regression_points <- get_regression_points(lifeExp_model, ID = "country")
regression_points
```

```{r LC5.6}
#5 countries with the smallest residuals
regression_points %>% 
  arrange(desc(residual)) %>% 
  top_n(5)

#5 countries with the largest residuals
regression_points %>% 
  arrange(residual) %>% 
  top_n(-5)

```

```{r Best fitting Line}
# Fit regression model:
score_model <- lm(score ~ bty_avg, 
                  data = evals_ch5)

# Get regression points:
regression_points <- get_regression_points(score_model)
regression_points

# Compute sum of squared residuals
regression_points %>%
  mutate(squared_residuals = residual^2) %>%
  summarize(sum_of_squared_residuals = sum(squared_residuals))
```

